SEDONA: A Novel Protocol for Identifying Infrequent, Long-Running Daemons on a Linux System

### 2017-01-25 sedona-ieice16/submission/paper.tex
sedona paper added

\documentclass[letter]{ieice}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{amssymb,amsthm,amsmath}
\usepackage{balance}
\usepackage{graphicx}          % graphicx package for including ps files  
\usepackage{epsfig}
\usepackage{url}
\usepackage{subfigure}
\usepackage{algorithmic}
\usepackage{float}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\usepackage{latexsym}
\usepackage{multirow}
\usepackage{color}
\usepackage{url}
\usepackage{comment}
%\usepackage{pstricks}
\usepackage{epstopdf}

@@

\field{D}
%\SpecialIssue{}
%\SpecialSection{}
%\theme{}
\title{SEDONA: A Novel Protocol for Identifying Infrequent, Long-Running Daemons on a Linux System}
%\titlenote{This work was in part supported by the EDISON program (NRF-2011-0020576).}
%\titlenote{}
\authorlist{
 \authorentry[yksuh@kisti.re.kr]{Young-Kyoon SUH}{m}{labelA}
 %\authorentry[rts@cs.arizona.edu]{Richard T. Snodgrass}{n}{labelB}}
\affiliate[labelA]{Supercomputing R\&{D} Center, KISTI, 245 Daehangno, Yuseong-gu, Daejeon, 34141, Republic of Korea}
%\affiliate[labelB]{Department of Computer Science, University of Arizona, Tucson, AZ 85721, USA}
}
\received{2016}{11}{30}
%\revised{2015}{10}{31}

@@

\begin{summary}
It is challenging to accurately measure the execution-time of a given
program due to many extraneous factors. To enable better timing results
restrained, we present a novel protocol that identifies infrequent,
long-running daemons to enable effective elimination of executions
exhibiting these daemons.

@@

Consider a simple compute-bound program, termed {\em PUT} (program-under-test), 
as shown in Fig.~\ref{alg:put}. 
This program runs a nested for-loop with a specified task length ($tl$) (in seconds). 
The $tl$ value is translated to the corresponding 
number of iterations for which that for-loop is performed.

\vspace{-.25in}
\begin{figure}[h]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} PerformManyIncrements($tl$): \\
\STATE $t$ = TranslateToNumRepeats($tl$) \\
\FOR{$k$ = 1 to $t$ by 1}
	\FOR{$i$ = 1 to {UINT\_MAX}-1 by 1}
		\STATE $j$ += 1 \\
	\ENDFOR 
\ENDFOR 
\end{algorithmic}
\end{center}
\caption{Task Definition by a Program-Under-Test (PUT)\label{alg:put}}
\vspace{-.25in}
\end{figure}

\begin{figure}[t]
	\centering
	\subfigure[Initial Elapsed Time Measurement]{
		\includegraphics[width=0.32\textwidth]{128_sec_et_old_hist.eps}
		\label{fig:hist_before_et}
	}
	\subfigure[Program Time After Identified Daemon Removal]{
		\includegraphics[width=0.32\textwidth]{128_sec_pt_new_hist.eps}
		\label{fig:hist_after_pt}
	}
	\caption{An Example of Measurement Results on a \linebreak \hbox{Compute-bound} 
	Program-Under-Test~\label{fig:meas_comp}}
	\vspace{-.3in}
\end{figure}

We used 128~seconds as the task length ran the PUT program (designated as PUT128) 
800~times. Fig.~\ref{fig:meas_comp} shows two histograms of the timing results on the PUT128.
As can be seen in Fig.\ref{fig:hist_before_et}, 
most of the 800 executions measured in elapsed time, 
via  {\tt System.currentTimeMillis()} in Java, are gathered in the tallest bar 
of \hbox{128,000--129,000 msec,} with just a few executions outside of that bar. 
But these outliers seriously hurt the overall measurement quality: 
a standard deviation of 1,700 msec and a relative error of 0.03. 
\shorten{This result suggests adequately eliminating these outliers 
while retaining most of the measured executions.}

In contrast, our scheme, which we propose shortly, 
significantly reduces such variability.
The gist of the scheme is to (i)~focus on
{\em process time} (PT) rather than on elapsed time (ET) and 
(ii)~remove some executions that involve
{\em infrequent, \hbox{long-running} daemon processes}, which impact
the process time of the process. 
%(Typically, one tick is equivalent to 1/100 secs, or 1 {\tt HZ}.) 
Using PT is preferred, in that it considers 
the time taken for only a process of interest.
PT can be calculated as the sum of ticks (where one tick is equal to 10 msec)
in user and system mode, available via {\tt taskstats} C struct, 
provided by the Linux NetLink facility~\cite{Netlink}.
This use of PT and selective elimination 
results in improved 
measurement quality---in both standard deviation and relative error---
by three orders of magnitude, 
as illustrated in Fig.~\ref{fig:hist_after_pt}.

{\bf Related Work.} 
McGeoch introduced
two basic methods of measuring program time: elapsed time and CPU time~\cite{Mcgeoch12}. 
Bryant and O'Hallaron~\cite{Randal03} 
presented two timing schemes of using clock-cycle and interval counters. 
They proposed a measurement protocol, called {\em minimum-of-k}, 
that for observed elapsed ticks the minimum is chosen as the most accurate one. 
Odom et al.'s work~\cite{Odom05} focuses on timing long-running programs 
in a simulation framework via 
dynamic sampling of trace snippets during program execution. 
None of these prior works, however, takes into account the variability in 
timing and the influence of daemons that may significantly disturb the timing.

%Some prior works~\cite{Burns93,Grund11,Santos10,Wilhelm08} discuss 
%the worst case execution time (WCET) of a process. 
%The papers focused on measuring the WCET of a deadline-sensitive process 
%on real-time embedded systems while being oblivious to 
%daemons coexisting with the process. In addition, we do not consider real-time constraints.

@@

{\bf Contribution.} 
Our contributions are following.
\vspace{-0.05in}
\begin{itemize}

\item We provide empirical evidence that 
measuring program time 
can be seriously affected by extant system daemons.

\item We propose a novel timing protocol that
identifies infrequent, long-running daemons that impact the timing results for that program. 
%This scheme determines in a disciplined way cutoffs to remove executions with such daemons.

@@

%\item We examine how the FIND scheme applies to process time as well as elapsed time.

\item The experimental results show a demonstrate for the effectiveness of our scheme\shorten{SEDONA: 
early eliminating dirty executions infected with such a daemon 
and eventually contributing to enhancing the overall measurement-quality}.

@@

In this section we propose a novel execution-time measurement scheme, 
called {\em SEDONA} (Selective Elimination through Detection of infrequent, lOng-ruNning dAemons). 
Our scheme catches and eliminates executions including daemon processes that are infrequent and 
long-running via a {\em cutoff} measure, and significantly improves measurement quality. 
The SEDONA scheme consists of a total of ten steps, as described in Fig.~\ref{alg:find}. 
%We first enumerate the steps to (a)~identify infrequent, 
%long-running daemons, using a single run of many executions of PUT128, 
%(b)~refine the list using a single run of many executions of PUT16384 (with a task length of 16K sec), 
%and then (c)~use the data from those two runs to determine the cutoff for 
%each \hbox{so-identified} daemon. 
%We can then apply these cutoffs to remove outliers from subsequent runs of any 
%arbitrary PUT and realistic workloads.
%\vspace{-0.22in}

\begin{figure}[h]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} The SEDONA Timing Scheme: \\
\STATE Step 1. Set up the timing environment.
\STATE Step 2. Perform a single PUT run (specifically, PUT128) for many samples (specifically, 800).
\STATE Step 3. Consider each pair of elapsed time measurements to be a
dual-PUT measurement 
and examine a scatter-plot to see if it it displays an $L$-shape.
\STATE Step 4. Zoom into the central cluster to ensure that it is symmetric (roughly circular).
\STATE Step 5. Compute the maximum and standard deviation of the process time 
for each daemon encountered within the central cluster samples.
\STATE Step 6. Identify for each sample in the $L$-shape infrequent, long-running daemon executions. 
\STATE Step 7. Determine potentially periodic daemons based on the $L$-executions 
and for each daemon compute the minimum process time from those executions identified. 
\STATE Step 8. Perform Steps 1--6 above for a single run 
consisting of a small number of executions (specifically, 40) of PUT16384.  
\STATE Step 9. Compute the cutoffs for each identified daemon. 
\STATE Step 10. Discard an execution including a daemon of which process
time is greater than the respective cutoff time.
\end{algorithmic}
\end{center}
\caption{Summary of the SEDONA Scheme\label{alg:find}}
\vspace{-0.28in}
\end{figure}
%

@@

\begin{figure}[t]
	\centering
	\subfigure[PUT128 with 800 Samples]{
		%\includegraphics[width=0.223\textwidth]{put128_empv4.eps}
		\includegraphics[width=0.223\textwidth]{put128_empv4_new.eps}
		\label{fig:reg_put128}
	}
	\subfigure[Dual-PUT256 Measurements]{
		%\includegraphics[width=0.223\textwidth]{dual_et_put256_empv4.eps}
		\includegraphics[width=0.223\textwidth]{dual_et_put256_empv4_new.eps}
		\label{fig:raw_put128}
	}
	%\vspace{-0.1in}    
	\subfigure[Zooming in on the $L$-shape]{
		\includegraphics[width=0.223\textwidth]{zooming_in_dual_et_put256_empv4.eps}
		\label{fig:dual_put256_liters}
	}
%	\subfigure[Further Zooming in on the $L$-Shape]{
%		\includegraphics[scale=0.39]{further_zooming_in_dual_et_put256_empv4.eps}
%		\label{fig:fzoom_in_dual_et_put256}
%	}
	\subfigure[The Central Cluster]{
		\includegraphics[width=0.223\textwidth]{dual_et_clump_put256_empv4.eps}
		\label{fig:dual_et_put256_cluster}
        }
    \caption{Successive Scatter plots of a PUT128 with 800 samples 
    (equivalent to a Dual-PUT256 with 400 samples) in Steps~2---4~\label{fig:put128_plot}} 
    \vspace{-0.3in}     
\end{figure}

@@

In Step 8 we repeat Steps 1--6, but instead with the much-longer running
PUT16384 (4.5 hours per sample versus 2 minutes), to see 
if any of our identified infrequent daemons are actually frequent 
at that much longer PUT execution time. 
We find some frequent daemon processes appearing in both of the 
clusters of the dual-PUT256 and dual-PUT32768: 
{\tt flush-9.0}, {\tt jbd2/md0-8}, {\tt kblock/0}, {\tt md0\_raid1}, 
and {\tt ntpd}. That said, the central cluster also contains other
  processes not seen in the dual-PUT256 central cluster: 
  {\tt grep}, {\tt rhn\_check}, {\tt rhnsd}, {\tt rhsmcertd}, 
  {\tt rhsmcertd-worke}, and {\tt sshd}. 
  But these daemons were categorized in the \hbox{dual-PUT256} analysis as {\em
  infrequent}, several having periodicities estimated at four or twenty hours.
When the PUT had a ``short'' 
program time (in this case, two minutes), daemons with a periodicity of
hours are infrequent. But with a PUT with a ``long'' program time (in this
case, 4.5 hours), some of those daemons are now frequent, and appear in the 
central cluster.

In Step~9 we compute 
the cutoff for each of those infrequent, long-running daemons so
identified, based on the PUT128 and the 
PUT16384 as collected in Tab.~\ref{tab:final_infrequent_cutoff}. 
Here is how to compute the cutoff. For the cutoff of such a daemon with PUT128, 
we take the midpoint between the maximum of that daemon's PTs in the 
central cluster (or 0, if absent) and the minimum of those in the $L$-samples. 
For the cutoff of such a daemon with PUT16384, we do the same. 
We then compute a ``task time'' as 5\% of the inferred periodicity. 
This 5\% ensures that such infrequent daemons will impact 
only a small percentage of the shorter PUTs, while presumably being 
associated with much larger cutoffs for the very long PUTs. 
We also include daemons that (a)~were identified as infrequent and 
\hbox{long-running} from PUT128 and (b)~were not identified as so in the PUT16384 
\hbox{$L$-samples}, but may have in the \hbox{dual-PUT32768} central cluster. 
We then take the {\em maximum} of the two cutoffs for the final cutoff PT 
(the last column of Tab.~\ref{tab:final_infrequent_cutoff}).

\begin{table}[h]
\centering
{\scriptsize
\begin{tabular}{|p{1.2cm}|c|c|p{1cm}|p{1.45cm}|} \hline
Process  & Cutoff PT & Cutoff PT & Task  & {\bf Final } \\
 Name & on PUT128  & on PUT16K & Time & {\bf Cutoff PT} \\\hline
{\tt bash} & 1 msec & --- & --- & {\bf 1 msec} \\ \hline
{\tt flush-9:0} & 64 msec & --- & $< 1$ hour & {\bf 64 msec} \\
                & ---     & 48 msec & $\geq 1$ hour & {\bf 48 msec} \\ \hline
{\tt grep }     & 1 msec & 12 msec & --- & {\bf 12 msec}\\ \hline
{\tt jbd2/md0-8} & 4 msec & --- & $< 1$ hour & {\bf 4 msec} \\
                & ---     & 11 msec & $\geq 1$ hour & {\bf 11 msec} \\ \hline
{\tt md0\_raid1} & 35 msec & ---     & $< 1$ hour  & {\bf 35 msec}\\
                 & ---     & 51 msec & $\geq 1$ hour & {\bf 51 msec} \\ \hline
{\tt rhn\_check}  & 281 msec & --- & $< 12$ min & {\bf 281 msec} \\
                 &  --- & 12,828 msec & $\geq 12$ min & {\bf 12,828 msec}\\ \hline
{\tt rhnsd} & 2 msec & --- & $< 12$ min & {\bf 2 msec} \\
            & ---    & 12 msec &$\geq 12$ min & {\bf 12 msec} \\ \hline
{\tt rhsmcertd}  & 1 msec & 1 msec & --- & {\bf 1 msec} \\  \hline
{\tt rhsmcertd}  & 57 msec & --- & $< 12$ min & {\bf 57 msec} \\
{\tt -worke}           &  --- & 119 msec  & $\geq 12$ min & {\bf 119 msec}\\ \hline
{\tt sshd} & 2 msec & 23 msec & --- & {\bf 23 msec}\\ \hline
\end{tabular}
}
\caption{Collected Infrequent, Long-running Daemons and Their Final Cutoff
  Process Time \hbox{(Step~9)}\label{tab:final_infrequent_cutoff}}
  \vspace{-0.4in}
\end{table}

Based on Tab.~\ref{tab:final_infrequent_cutoff}, 
we discard any sample containing an infrequent, long-running daemon execution
over the corresponding cutoff. We thus end up dropping just fifteen of the 800
PUT128 samples and only two of the forty PUT16384 samples. 
As a result, the overall measurement quality---the standard deviation and relative error 
---for PUT128 was improved 
by three orders of magnitude (as shown in Fig.~\ref{fig:meas_comp}) 
and by about two orders of magnitude for PUT16384.
%(5661.7 vs. 229.8) 0.0003448491 vs. 8.4$\times$10$^{-6}$
%(40.4 vs. 26.7) 2.463122e-06 vs. 1.626005e-06

%\begin{table}[H]
%\centering
%{\small
%\begin{tabular}{|c|c|c|} \hline
% & \multirow{2}{*}{ET/PT} & Std. Dev. & \multirow{2}{*}{Rel. Err.} \\ 
% & & &  & (msec) &  \\ \hline
%\multirow{2}{*}{PUT128} & ET & 800 & 2.5  & 2.0$\times$10$^{-5}$\\  \cline{2-9}
%  & PT & 2.6  & 2.0$\times$10$^{-5}$\\ \hline  
%\multirow{2}{*}{PUT16384} & ET & 138.6  & 8.4$\times$10$^{-6}$\\  \cline{2-9}
% & PT & 2.7  & 1.6$\times$10$^{-6}$\\ \hline  
%\end{tabular}
%}
%\caption{Measurement Statistics of PUT128 (above) and PUT16384 (below) 
%by the SEDONA Scheme\label{tab:refined_stat}}
%\end{table}

%We additionally remove those samples 
%in which the PT measurements are greater or less 
%two standard deviations from the average of the combined samples. In this
%particular case, no additional samples were removed.

%Finally, we calculate the execution time of PUT128 as the average PT measurement in Table~\ref{tab:refined_empv5}: 
%128250 milliseconds (rounded up to the nearest millisecond, because PT is
%measured in msec). Table~\ref{tab:refined_empv5} shows the statistics of 
%the PT and ET measurements. (In the fourth column, the number of samples
%dropped by the first sanity check (15 for ET) and by the second sanity check
%(1) are indicated.) Our original goal was to
%determine using a carefully-motivated protocol the program time (PT) for
%this program under test (PUT). We see that the standard deviation and
%relative error have both been considerably reduced by this principled
%elimination of samples containing identified infrequent, long-running daemons.

>>>>

\documentclass[letter]{ieice}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{amssymb,amsthm,amsmath}
\usepackage{balance}
\usepackage{graphicx}          % graphicx package for including ps files  
\usepackage{epsfig}
\usepackage{url}
\usepackage{subfigure}
\usepackage{algorithmic}
\usepackage{float}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\usepackage{latexsym}
\usepackage{multirow}
\usepackage{color}
\usepackage{url}
\usepackage{comment}
%\usepackage{pstricks}
\usepackage{array}
\usepackage{epstopdf}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

\field{D}
\title{SEDONA: A Novel Protocol for Identifying Infrequent, Long-Running Daemons on a Linux System}
%\titlenote{This work was in part supported by the EDISON program (NRF-2011-0020576).}
\authorlist{
 \authorentry[yksuh@kisti.re.kr]{Young-Kyoon SUH}{m}{labelA}
 %\authorentry[rts@cs.arizona.edu]{Richard T. Snodgrass}{n}{labelB}}
\affiliate[labelA]{Supercomputing R\&{D} Center, KISTI, 245 Daehangno, Yuseong-gu, Daejeon, 34141, Republic of Korea}
%\affiliate[labelB]{Department of Computer Science, University of Arizona, Tucson, AZ 85721, USA}
}
\received{2016}{12}{3}

@@

\begin{summary}
It is challenging to accurately measure the execution-time of a given
program due to many extraneous factors. To enable better timing results\shorten{restrained},
we present a novel protocol that identifies infrequent, 
long-running daemons to enable effective elimination of executions 
exhibiting these daemons.

@@

Consider a simple compute-bound program, what we term {\em PUT} (program-under-test), 
as shown in Fig.~\ref{alg:put}. 
This program runs a nested for-loop with a specified task length ($tl$) (in seconds). 
The $tl$ value is translated to the corresponding 
number of iterations for which that for-loop is performed.

\vspace{-.2in}
\begin{figure}[h]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} PerformManyIncrements($tl$): \\
\STATE $t$ = $tl$ * {CONSTANT} \\
\FOR{$k$ = 1 to $t$ by 1}
	\FOR{$i$ = 1 to {UINT\_MAX}-1 by 1}
		\STATE $j$ += 1 \\
	\ENDFOR 
\ENDFOR 
\end{algorithmic}
\end{center}
\caption{Computation by a Program-Under-Test (PUT)\label{alg:put}}
\vspace{-.2in}
\end{figure}

We used 128~seconds as the task length and ran the PUT program (termed PUT128) 
800~times. Fig.~\ref{fig:meas_comp} shows two histograms of the timing results on PUT128.
Most of the 800 executions measured in elapsed time, 
via  {\tt System.currentTimeMillis()} in Java, are gathered in the tallest bar 
of \hbox{128,000--129,000 msec,} with just a few executions outside of that bar, as can be seen in Fig.~\ref{fig:hist_before_et}. 
But these outliers significantly impact the overall measurement quality: 
a standard deviation of 1,700 msec and a relative error of 0.03. 
\shorten{This result suggests adequately eliminating these outliers 
while retaining most of the measured executions.}

In contrast, our scheme, which we propose shortly, 
significantly reduces such variability.
The gist of the scheme is to (i)~focus on
{\em process time} (PT) rather than on elapsed time (ET) and 
(ii)~remove some executions that involve
{\em infrequent, \hbox{long-running} daemon processes}, which impact
the process time of the process. 
Using PT is preferred, in that it considers 
the time taken for only a process of interest.
PT can be calculated as the sum of ticks (where one tick is equal to 10 msec)
in user and system mode, available via {\tt taskstats} C struct, 
provided by the Linux NetLink facility~\cite{Netlink}.
This use of PT and selective elimination 
was able to improve 
measurement quality---in both standard deviation and relative error---up to by 
three orders of magnitude for PUT128, as illustrated in Fig.~\ref{fig:hist_after_pt}.

%\vspace{0.05in}
{\bf Related Work.} 
McGeoch introduced
two basic methods of measuring program time: elapsed time and CPU time~\cite{Mcgeoch12}. 
Bryant and O'Hallaron~\cite{Randal03} 
presented two timing schemes of using clock-cycle and interval counters. 
They proposed a measurement protocol, called {\em minimum-of-k}, 
that for observed elapsed ticks the minimum is chosen as the most accurate one. 
Odom et al.'s work~\cite{Odom05} focused on timing long-running programs 
in a simulation framework via dynamic sampling of trace snippets during program execution. 
None of these prior works, however, takes into account the variability in 
timing and the influence of daemons that may significantly disturb the timing.

\begin{figure}[t]
	\centering
	\subfigure[Initial Elapsed Time Measurement]{
		\includegraphics[width=0.31\textwidth]{128_sec_et_old_hist.eps}
		\label{fig:hist_before_et}
	}
	\subfigure[Program Time After Identified Daemon Removal]{
		\includegraphics[width=0.31\textwidth]{128_sec_pt_new_hist.eps}
		\label{fig:hist_after_pt}
	}
	\caption{An Example of Measurement Results on a \linebreak \hbox{Compute-bound} 
	Program-Under-Test~\label{fig:meas_comp}}
	\vspace{-.18in}
\end{figure}

@@

%\vspace{0.05in}
{\bf Contribution.} 
Our contributions are following.
\vspace{-0.12in}
\begin{itemize}

\item We show empirical evidence that 
measuring program time 
can be seriously affected by daemons.

\item We propose a novel timing protocol that
identifies infrequent, long-running daemons that impact the timing results for that program.

@@

\item The experimental results show a support for the effectiveness of our scheme\shorten{SEDONA: 
early eliminating dirty executions infected with such a daemon 
and eventually contributing to enhancing the overall measurement-quality}.

@@

In this section we propose a novel execution-time measurement scheme, 
called {\em SEDONA} (Selective Elimination through Detection of infrequent, lOng-ruNning dAemons). 
Our scheme catches and eliminates executions including daemon processes that are infrequent and 
long-running via a {\em cutoff} measure, and significantly improves measurement quality. 
The SEDONA protocol consists of a total of ten steps, as described in Fig.~\ref{alg:find}. 
%We first enumerate the steps to (a)~identify infrequent, 
%long-running daemons, using a single run of many executions of PUT128, 
%(b)~refine the list using a single run of many executions of PUT16384 (with a task length of 16K sec), 
%and then (c)~use the data from those two runs to determine the cutoff for 
%each \hbox{so-identified} daemon. 
%We can then apply these cutoffs to remove outliers from subsequent runs of any 
%arbitrary PUT and realistic workloads.
%\vspace{-0.22in}

\begin{figure}[h]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} The SEDONA Timing Protocol: \\
\STATE Step 1. Set up the timing environment.
\STATE Step 2. Perform a single PUT run (specifically, PUT128) for many samples (specifically, 800).
\STATE Step 3. Consider each pair of elapsed time measurements to be a dual-PUT measurement 
and examine a scatter-plot to see if it it displays an $L$-shape.
\STATE Step 4. Zoom into the central cluster to ensure that it is symmetric (roughly circular).
\STATE Step 5. Compute the maximum and standard deviation of the process time 
for each daemon encountered within the central cluster samples.
\STATE Step 6. Identify for each sample in the $L$-shape infrequent, long-running daemon executions. 
\STATE Step 7. Determine potentially periodic daemons based on the $L$-executions 
and for each daemon compute the minimum process time from those executions identified. 
\STATE Step 8. Perform Steps 1--6 above for a single run 
consisting of a small number of executions (specifically, 40) of PUT16384.  
\STATE Step 9. Compute the cutoffs for each identified daemon. 
\STATE Step 10. Discard an execution including a daemon of which process time is greater than the respective cutoff time.
\end{algorithmic}
\end{center}
\caption{Summary of the SEDONA Timing Protocol\label{alg:find}}
\vspace{-0.25in}
\end{figure}
%

@@

\begin{figure}[t]
	\centering
	\subfigure[PUT128 with 800 Samples]{
		%\includegraphics[width=0.223\textwidth]{put128_empv4.eps}
		\includegraphics[width=0.224\textwidth]{put128_empv4_new.eps}
		\label{fig:reg_put128}
	}
	\subfigure[Dual-PUT256 Measurements]{
		%\includegraphics[width=0.223\textwidth]{dual_et_put256_empv4.eps}
		\includegraphics[width=0.224\textwidth]{dual_et_put256_empv4_new.eps}
		\label{fig:raw_put128}
	}
	\vspace{0.1in}    
	\subfigure[Zooming in on the $L$-shape]{
		\includegraphics[width=0.224\textwidth]{zooming_in_dual_et_put256_empv4.eps}
		\label{fig:dual_put256_liters}
	}
%	\subfigure[Further Zooming in on the $L$-Shape]{
%		\includegraphics[scale=0.39]{further_zooming_in_dual_et_put256_empv4.eps}
%		\label{fig:fzoom_in_dual_et_put256}
%	}
	\subfigure[The Central Cluster]{
		\includegraphics[width=0.224\textwidth]{dual_et_clump_put256_empv4.eps}
		\label{fig:dual_et_put256_cluster}
        }
    \caption{Successive Scatter plots of a PUT128 with 800 samples 
    (equivalent to a Dual-PUT256 with 400 samples) in Steps~2---4~\label{fig:put128_plot}} 
    \vspace{-0.16in}     
\end{figure}

@@

In Step 8 we repeat Steps 1--6, but instead with the much-longer running
PUT16384 (4.5 hours per \hbox{sample} versus 2 minutes), to see 
if any of our identified \hbox{infrequent} daemons are actually frequent 
at that much longer PUT execution time. 
We find some frequent daemon processes appearing in both of the 
clusters of the dual-PUT256 and dual-PUT32768: 
{\tt flush-9.0}, {\tt jbd2/md0-8}, {\tt kblock/0}, {\tt md0\_raid1}, 
and {\tt ntpd}. That said, the central cluster also contains other
  processes not seen in the dual-PUT256 central cluster: 
  {\tt grep}, {\tt rhn\_check}, {\tt rhnsd}, {\tt rhsmcertd}, 
  {\tt rhsmcertd-worke}, and {\tt sshd}. 
  But these daemons were categorized in the \hbox{dual-PUT256} analysis as {\em
  infrequent}, several having \hbox{periodicities} estimated at four or twenty hours.
When the PUT had a ``short'' 
program time (in this case, two minutes), daemons with a periodicity of
hours are \hbox{infrequent}. But with a PUT with a ``long'' program time (in this
case, 4.5 hours), some of those daemons are now frequent, and appear in the 
central cluster.

In Step~9 we compute 
the cutoff for each of those infrequent, long-running daemons so
identified, based on the PUT128 and the 
PUT16384 as collected in Tab.~\ref{tab:final_infrequent_cutoff}. 
Here is how to compute the cutoff. For the cutoff of such a daemon with PUT128, 
we take the \hbox{midpoint} between the maximum of that daemon's PTs in the 
central cluster (or 0, if absent) and the \hbox{minimum} of those in the $L$-samples. 
For the cutoff of such a daemon with PUT16384, we do the same. 
We then compute a ``task time'' as 5\% of the inferred periodicity. 
This 5\% ensures that such infrequent daemons will impact 
only a small percentage of the shorter PUTs, while presumably being 
associated with much larger cutoffs for the very long PUTs. 
We also include daemons that (a)~were identified as infrequent and 
\hbox{long-running} from PUT128 and (b)~were not identified as so in the PUT16384 
\hbox{$L$-samples}, but may have in the \hbox{dual-PUT32768} central cluster. 
We then take the {\em maximum} of the two cutoffs for the final cutoff PT 
(the last column of Tab.~\ref{tab:final_infrequent_cutoff}).

\begin{table}[h]
\centering
{\scriptsize
\begin{tabular}{|p{1.2cm}|c|c|p{1cm}|p{1.45cm}|} \hline
{\tiny Process}  & {\tiny Cutoff PT} & {\tiny Cutoff PT} & {\tiny Task}  & {\tiny{\bf Final }} \\
{\tiny Name} & {\tiny on PUT128}  & {\tiny on PUT16K} & {\tiny Time} & {\tiny {\bf Cutoff PT}} \\\hline
%{\tiny Process}  & {\tiny Cutoff PT} & {\tiny Cutoff PT} & {\tiny Task}  & {\tiny{\bf Final }} \\
% Name & on PUT128  & on PUT16K & Time & {\bf Cutoff PT} \\\hline
{\tt bash} & 1 msec & --- & --- & {\bf 1 msec} \\ \hline
{\tt flush-9:0} & 64 msec & --- & $< 1$ hour & {\bf 64 msec} \\
                & ---     & 48 msec & $\geq 1$ hour & {\bf 48 msec} \\ \hline
{\tt grep }     & 1 msec & 12 msec & --- & {\bf 12 msec}\\ \hline
{\tt jbd2/md0-8} & 4 msec & --- & $< 1$ hour & {\bf 4 msec} \\
                & ---     & 11 msec & $\geq 1$ hour & {\bf 11 msec} \\ \hline
{\tt md0\_raid1} & 35 msec & ---     & $< 1$ hour  & {\bf 35 msec}\\
                 & ---     & 51 msec & $\geq 1$ hour & {\bf 51 msec} \\ \hline
{\tt rhn\_check}  & 281 msec & --- & $< 12$ min & {\bf 281 msec} \\
                 &  --- & 12,828 msec & $\geq 12$ min & {\bf 12,828 msec}\\ \hline
{\tt rhnsd} & 2 msec & --- & $< 12$ min & {\bf 2 msec} \\
            & ---    & 12 msec &$\geq 12$ min & {\bf 12 msec} \\ \hline
{\tt rhsmcertd}  & 1 msec & 1 msec & --- & {\bf 1 msec} \\  \hline
{\tt rhsmcertd}  & 57 msec & --- & $< 12$ min & {\bf 57 msec} \\
{\tt -worke}           &  --- & 119 msec  & $\geq 12$ min & {\bf 119 msec}\\ \hline
{\tt sshd} & 2 msec & 23 msec & --- & {\bf 23 msec}\\ \hline
\end{tabular}
}
\caption{Collected Infrequent, Long-running Daemons and Their Final Cutoff
  Process Time \hbox{(Step~9)}\label{tab:final_infrequent_cutoff}}
  \vspace{-0.4in}
\end{table}

Based on Tab.~\ref{tab:final_infrequent_cutoff}, 
we discard any \hbox{sample} \hbox{containing} an infrequent, long-running daemon execution
over the corresponding cutoff. We thus end up dropping just fifteen of the 800
PUT128 samples and only two of the forty PUT16384 samples. 
As a result, the \hbox{overall} \hbox{measurement} quality---the standard deviation and \hbox{relative} error 
---for PUT128 was improved 
by three orders of magnitude (as shown in Fig.~\ref{fig:meas_comp}) 
and by about two \hbox{orders} of magnitude for PUT16384.
%(5661.7 vs. 229.8) 0.0003448491 vs. 8.4$\times$10$^{-6}$
%(40.4 vs. 26.7) 2.463122e-06 vs. 1.626005e-06

\vspace\fill

\section{Evaluation}
\label{sec:eval}
\vspace{-0.07in}
We now evaluate the \hbox{performance} of the SEDONA \hbox{protocol}.
Our experiments were conducted on a \hbox{machine}
described in Table~\ref{tab:machine_config}. 
\begin{table}[h]
\vspace{-0.2in}
\begin{center}
{\scriptsize
\begin{tabular}{|l|p{7cm}|}\hline
OS & Red Hat Ent. Linux (RHEL) 6.4 with a kernel of 2.6.32 \\ \hline
CPU & Intel Core i7-870 Lynnfield 2.93GHz quad-core \hbox{processor}\shorten{ on a LGA 1156 95W motherboard}\\ \hline
RAM & 4GB of DDR3 1333 dual-channel memory\\ \hline
HDD & Western Digital Caviar Black 1TB 7200rpm SATA Drive\\ \hline
\end{tabular}
}
\end{center}
\caption{Machine Configurations\label{tab:machine_config}}
\vspace{-0.3in}
\end{table}

We evaluated the performance of SEDONA \hbox{using} SPEC CPU2006 
benchmarks~\cite{specCpu2006}, providing \hbox{various} compute-bound real applications. 
%Thus, employing the SPEC benchmarks as real-world workloads 
%was a reasonable choice for the evaluation. 
The results are \hbox{provided} in Table~\ref{tab:spec_real}. 
Note that in the table the results for {\tt 481} and {\tt 483} \hbox{benchmarks} 
are omitted because of some runtime error and incurred I/O, respectively.

Table~\ref{tab:spec_real} shows that
the SEDONA protocol (which uses PT) significantly outperformed the original 
measurement technique (using ET), termed ORG, 
on the standard deviation and relative error across the different SPEC benchmarks. 
None of the benchmarks revealed a bigger standard deviation from the SEDONA protocol as compared to that of ORG. 
%All of the benchmarks revealed a smaller standard deviation on our scheme than that of the ORG. 
Our protocol quite effectively filtered out infrequent daemon executions 
in these real-world workloads. 
Furthermore, the relative error of SEDONA was 
lower than that of ORG for all the benchmarks. 
For instance, about a 10x \hbox{margin} between the two was yielded for 
{\tt 434.zeusmp}.

SEDONA also scaled well for the SPEC workloads, 
with regard to growth of relative error as the execution time lengthened.
For the short benchmarks 
(e.g., {\tt 400}, {\tt 403}, {\tt 410}, 
{\tt 434}, {\tt 445}, and {\tt 999}: those taking \hbox{under} 100 sec), 
our scheme outperformed the ORG by about 3.5x, on average. 
The SEDONA protocol continued its dominance against the conventional technique 
for the medium-length benchmarks (e.g., {\tt 447}, {\tt 456}, {\tt 470}, and {\tt 473}).
For the long-running benchmarks (e.g., {\tt 436} and {\tt 454}, both$>$ 1,000 sec), 
the relative error of \hbox{SEDONA} was slightly lower than that of ORG.

%Lastly, the average standard deviation and relative error of the SEDONA 
%among all the benchmarks were smaller that those of the ORG, 
%as indicated by the last two rows. 
%Lastly, the SEDONA outperformed the ORG by a slight margin \hbox{regarding} 
%the plain (un-weighted) and weighted averages of 
%the standard deviation and relative error among all the benchmarks, 
%as indicated by the last two rows.
%\shorten{In particular, the SEDONA outperformed the ORG regarding 
%the weighted average by normalization by a slight margin 
%in spite of seemingly being equal because of the rounding up to one significant digit.}
